---
title: "PCSF Tests"
author: "Sara Gosline"
date: "5/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(PCSF)
library(ggplot2)
library(dplyr)
library(tidyr)
setwd('..')
source("loadHumanPancData.R")
```

## Building trees out of voxels

The idea behind this approach is to build steiner trees out of individual voxels. In doing so we can make the voxels more functionally relevant. 

```{r map ids}



up_map <- read.table("../uniprotMap.txt",col.names=c('protein','genename'))

fulltab<-crosstab2%>%
  as.data.frame()%>%
  tibble::rownames_to_column('geneid')%>%
  tidyr::separate(geneid,into=c('id','upid','protein'),
                  sep='\\|')%>%
  tidyr::pivot_longer(contains('_S_'),
                      names_to='spot',
                      values_to='logRatio')%>%
  #unite(islet,Coords,sep='_',col='spot',remove = FALSE)%>%
  left_join(tibble::rownames_to_column(isletMeta,'spot'))%>%
  left_join(up_map)

```



## Get most expressed proteins in each voxel

First we bracket the data into the top vs bottom proteins to see how they are expressed.


```{r get top proteins}

##group by islet, take top 2.5 and bottom 2.5% of proteins
  
limits <- fulltab%>%
  dplyr::select(spot,logRatio)%>%
  dplyr::group_by(spot)%>%
  summarize(top=quantile(logRatio,0.95,na.rm=T),bottom=quantile(logRatio,0.05,na.rm=T))

##now determine if proteins are in top or bottom or none
topTerms <- fulltab%>%left_join(limits)%>%
  mutate(isTop=logRatio>top)%>%
  mutate(isBottom=logRatio<bottom)%>%
  subset(isTop|isBottom)

##how many top proteins in each spot? plot this
ggplot(topTerms)+geom_jitter(aes(x=spot,y=logRatio,col=IsletStatus))

protCounts<-fulltab%>%group_by(spot)%>%summarize(totalProts=n_distinct(genename))
topCounts<-topTerms%>%group_by(spot,IsletStatus)%>%summarize(totalTop=n_distinct(genename))%>%full_join(protCounts)
  

```

##Now run networks on top prots

iterate over the top proteins in each voxel 

```{r run net, echo=FALSE, warning=FALSE,message=FALSE}


library(PCSF)
data('STRING')
runPcsfOnProts<-function(tab,spot,nrand=10){
  fname=paste0('spot',spot,'nrand',nrand)
  terms<-abs(tab$logRatio)
  spot=unlist(spot)
#  print(spot)
  gene<-tab$genename
  ##map protein to gene name
  names(terms)<-gene
  ppi <- construct_interactome(STRING)
  
  ##return gene names, prizes
  print(spot)
  nres<-data.frame(nodes=c(""),prizes=c(""),spot=spot)
  res<-NULL
  try(res<-PCSF::PCSF_rand(ppi=ppi,terminals = terms,n=nrand))
  if(!is.null(res) && length(vertex.attributes(res)$name)>1){
    write_graph(res,format='gml',file=paste0(fname,'.gml'))

    nres<-data.frame(nodes=vertex.attributes(res)$name,
                     prizes=vertex.attributes(res)$prize,
                     spot=spot)
  }
  return(nres)
}

##build network for each voxel

#now build networks based on top terms alone 
nets<-topTerms%>%
  subset(isTop)%>%
  dplyr::select(spot,genename,logRatio)%>%
 # separate(protein,into=c('gene','human'),sep='_')%>%
#  dplyr::select(spot,genename,logRatio)%>%
  group_by(spot)%>%
  dplyr::group_map(.f=runPcsfOnProts,.keep=FALSE)%>%
  do.call(rbind,.)

##now join with data from original table
##map distance between each network?
toplot<-fulltab%>%dplyr::select(genename,spot,logRatio,IsletStatus)%>%
  #tidyr::separate(protein,into=c('nodes','hum'),sep='_')%>%
  dplyr::rename(nodes='genename')%>%
  right_join(nets,by=c('nodes','spot'))
##for each list of proteins in each voxel network, identify overlap with expresion data


net_counts<-nets%>%
  group_by(spot)%>%
  summarize(netProts=n_distinct(nodes))%>%
  full_join(topCounts)%>%
  tidyr::pivot_longer(cols=c(netProts,totalTop,totalProts),names_to='group',values_to='Number proteins')

ggplot(net_counts,aes(x=IsletStatus,y=`Number proteins`,col=group))+geom_jitter()

```


This is kind of a bummer, as the networks appear to just be reducing the protein set.

## then plot PCA of new reduced expression set

Plotting the PCA values of the proteins in the networks doesn't show much difference - whether we use prizes or expression.

```{r pca vals}

##PCA of prizes
df<-toplot%>%dplyr::select(spot,prizes,nodes)%>%
  mutate(prizes=as.numeric(prizes))%>%
  distinct()%>%
  tidyr::pivot_wider(names_from=spot,values_from = prizes,
                     values_fill = list(prizes=0.0),
                     values_fn = list(prizes=mean))%>%
  tibble::column_to_rownames('nodes')

mtab<-toplot%>%
  dplyr::select(spot,IsletStatus)%>%
  distinct()%>%
  subset(!is.na(IsletStatus))%>%
  tibble::column_to_rownames('spot')

nac<-which(colMeans(df,na.rm=T)!=0)
nar<-which(rowMeans(df,na.rm=T)!=0)
pca_res <- prcomp(t(df[nar,nac]),scale.=TRUE,center = TRUE)

p<-autoplot(pca_res,data=mtab,colour='IsletStatus')
print(p)


##PCA of expression
df2<-toplot%>%dplyr::select(spot,logRatio,nodes)%>%
  mutate(logRatio=as.numeric(logRatio))%>%
  distinct()%>%
  replace_na(list(logRatio=0.0))%>%
  tidyr::pivot_wider(names_from=spot,
                     values_from = logRatio,
                     values_fn =list(logRatio=mean),
                     values_fill = list(logRatio=0.0))%>%
  tibble::column_to_rownames('nodes')

mtab2<-toplot%>%dplyr::select(spot,Islet_status)%>%
  distinct()%>%
  subset(!is.na(Islet_status))%>%
  tibble::column_to_rownames('spot')

nac<-which(colMeans(df2,na.rm=T)!=0)
nar<-which(rowMeans(df2,na.rm=T)!=0)
pca_res <- prcomp(t(df2[nar,nac]),scale.=TRUE,center = TRUE)

p<-autoplot(pca_res,data=mtab,colour='Islet_status')
print(p)

```


## Use all proteins

What if we just throw all proteins onto the network? will that help 'fill in the gap'?

```{r big networks}



#now build networks based on top terms alone 
#bigsnets<-fulltab%>%
#  dplyr::select(spot,genename,logRatio)%>%
#  mutate(nrand=10)%>%
 # separate(protein,into=c('gene','human'),sep='_')%>%
#  dplyr::select(spot,genename,logRatio)%>%
#  group_by(spot)%>%
#  dplyr::group_map(.f=runPcsfOnProts,.keep=FALSE)%>%
#  do.call(rbind,.)



##PCA of prizes
#df<-bigsnets%>%dplyr::select(spot,prizes,nodes)%>%
#  mutate(prizes=as.numeric(prizes))%>%
#  distinct()%>%
#  tidyr::pivot_wider(names_from=spot,values_from = prizes,
#                     values_fill = list(prizes=0.0),
#                     values_fn = list(prizes=mean))%>%
#  tibble::column_to_rownames('nodes')

#mtab<-toplot%>%
#  dplyr::select(spot,Islet_status)%>%
#  distinct()%>%
#  subset(!is.na(Islet_status))%>%
#  tibble::column_to_rownames('spot')

# nac<-which(colMeans(df,na.rm=T)!=0)
# nar<-which(rowMeans(df,na.rm=T)!=0)
# pca_res <- prcomp(t(df[nar,nac]),scale.=TRUE,center = TRUE)
# 
# p<-autoplot(pca_res,data=mtab,colour='Islet_status')
# print(p)
# 
# 
# ##PCA of expression
# df2<-toplot%>%dplyr::select(spot,logRatio,nodes)%>%
#   mutate(logRatio=as.numeric(logRatio))%>%
#   distinct()%>%
#   replace_na(list(logRatio=0.0))%>%
#   tidyr::pivot_wider(names_from=spot,
#                      values_from = logRatio,
#                      values_fn =list(logRatio=mean),
#                      values_fill = list(logRatio=0.0))%>%
#   tibble::column_to_rownames('nodes')
# 
# mtab2<-toplot%>%dplyr::select(spot,Islet_status)%>%
#   distinct()%>%
#   subset(!is.na(Islet_status))%>%
#   tibble::column_to_rownames('spot')
# 
# nac<-which(colMeans(df2,na.rm=T)!=0)
# nar<-which(rowMeans(df2,na.rm=T)!=0)
# pca_res <- prcomp(t(df2[nar,nac]),scale.=TRUE,center = TRUE)
# 
# p<-autoplot(pca_res,data=mtab,colour='Islet_status')
# print(p)
```


## Network diffusion?

Vince will do this..
```{R diffuse network}



##let's borrow a bit from PCSF
modstring <- STRING
modstring$cost <- 1-STRING$cost
igraph_ppi <- construct_interactome(modstring)

##https://github.com/USCCANA/netdiffuseR
#or 
###https://www.bioconductor.org/packages/devel/bioc/vignettes/diffuStats/inst/doc/diffuStats.pdf
```
## Now compare trees in PCA to see how they compare



## can we use tree output as input? 
