---
title: "HubMap pancreata exploratory figures and statistical analysis"
output: html_document
---

This document describes the basic analysis we need to showcase the value of the Hubmap data and proteomics analysis.

```{r loading, echo=F, warning=F, message=F}
library(dplyr)
library(purrr)
library(devtools)
require(NatParksPalettes)
#library(MSnSet.utils)
if(!require(MSnSet.utils))
  devtools::install_github('PNNL-Comp-Mass-Spec/MSnSet.utils')

library(ggplot2)
library(ggfortify)
library(cowplot)
#library(ggbiplot)

source('loadHumanPancData.R')

```

Note that we store all files on synapse at [http://synapse.org/hubmap](http://synapse.org/hubmap). This includes basic data files and analysis results.

Now we have a few ways to normalize.

### First we have to test the effect of normalization

Let's plot the distribution of all of the values, test for normality, and plot the PCs.


```{r normality test}
library(ggridges)
library(NatParksPalettes)
p1<-ggplot(dplyr::mutate(fulltab,Grid=as.factor(`Grid Number`)),
       aes(x=logRatio,fill=Image,y=Grid))+
  geom_density_ridges2(alpha=0.5)+
  scale_fill_manual(values=natparks.pals('RockyMtn',7))+
  theme_bw()

p1

ggsave('a_countdistributionPlot.pdf',p1)


# #shap_ps<-fulltab%>%
#     group_by(Spot)%>%
#     dplyr::summarise(norm_p=shapiro.test(sample(logRatio,size=4400))$p.value)%>%
#   mutate(logP=log10(norm_p))
# 
# ggplot(dplyr::mutate(normtab,Grid=as.factor(`Grid Number`)),
#        aes(x=logRatio,fill=Grid,y=Image))+
#   geom_density_ridges2(alpha=0.5)
# 
# shap_norm_ps<-normtab%>%
#   as.data.frame()%>%
#   group_by(Spot)%>%
#   dplyr::summarise(norm_p=shapiro.test(sample(logRatio,size=4400))$p.value)

```

Now that we have visualized the data, we need to count the number of proteins, those with complete coverage, and how many overall NA values we have.  For the purposes of this analysis, we set the NA values to the median for each image. 

```{r account for missing data}
##here we plot number of proteins missing across each image?

missing<-fulltab%>%
  dplyr::mutate(Grid=as.factor(`Grid Number`))%>%
  dplyr::group_by(Spot,Image,Grid,IsletOrNot)%>%
  dplyr::summarize(count=n(),nas=sum(is.na(logRatio)))%>%
  mutate(fracNAs=nas/count)

p2<-ggplot(missing,aes(x=Grid,y=fracNAs,fill=Image))+
  geom_bar(stat='identity',position='dodge')+
   scale_fill_manual(values=natparks.pals('RockyMtn',7))+
  theme_bw()
  

ggsave('b_missingProts.pdf',p2)
p2

ggplot(missing,aes(x=Image,y=nas,fill=IsletOrNot))+geom_bar(stat='identity',position='dodge')

```

```{r PCA plots}

fixed.crosstab<-correctMissingProteins(fulltab)


##plot original components

##plot 1 vs. 2, and 2 vs. 3



pc.z<-prcomp(t(fixed.crosstab), center=TRUE,scale.=TRUE)
#pc.red<-prcomp(t(red.crosstab), center=TRUE,scale.=TRUE)


#autoplot(pc,data=isletMeta,colour='IsletNumber')
plot(pc.z)

p3<-autoplot(pc.z,data=isletMeta,colour='IsletStatus')+
  scale_color_manual(values=natparks.pals('Saguaro',3))+
  theme_bw()

#p3<-autoplot(pc.z,data=isletMeta,colour='IsletStatus')+ggtitle('Full data')

p4<-autoplot(pc.z,data=isletMeta,colour='IsletNumber')+
  scale_color_manual(values=natparks.pals('RockyMtn',7))+
  theme_bw()

res<-cowplot::plot_grid(p1,p2,p3,p4,nrow = 2,ncol=2)
res

ggsave('fig2_normalization.pdf',res)
# 
# ##now plotting  with msn code
# isletMeta <- isletMeta[colnames(crosstab), ]
# 
# ##we keep NA values in m1 and m2
# red.crosstab <- crosstab[rowMeans(!is.na(crosstab)) >= .5, ]
# 
# m1 <- MSnSet(exprs = red.crosstab, pData = isletMeta)
# 
# m2<- MSnSet(exprs = crosstab, pData = isletMeta)
# ##Need to find the plot_pca funciton, is it in msnbase.utils?
# #plot_pca(m1, phenotype = "Islet_Number") #no batch effect
# p3=plot_pca(m1, phenotype = "IsletStatus",z_score=T)+ggtitle('Reduced data') #check biplot = T?
# p4=plot_pca(m2, phenotype = "IsletStatus",z_score=T)+ggtitle("Full data") #check biplot = T?
# 
# cowplot::plot_grid(p3,p4)
# 
# p5=plot_pca(m1, phenotype = "IsletNumber",z_score=T)+ggtitle('Reduced data') #check biplot = T?
# p6=plot_pca(m2, phenotype = "IsletNumber",z_score=T)+ggtitle("Full data") #check biplot = T?
# 
# cowplot::plot_grid(p5,p6)
#save(m1, file = "combinedIslets_Results_New/combinedMSnSet.RData", compress = T)


```



