---
title: "Assessing the predictive accuracy of islets from data"
author: "Sara Gosline"
date: "5/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(BayesSpace)
library(ggplot2)
library(dplyr)
library(tidyr)

  library(RColorBrewer)
  library(cowplot)
  
```

## Format data
First grab data and format into single table. Evaluate distributions of values. There are clearly different distributions when we median center the data.

```{r get data, message=F,warning=F}
library(dplyr)
library(purrr)
library(devtools)
#library(MSnSet.utils)
if(!require(MSnSet.utils))
  devtools::install_github('PNNL-Comp-Mass-Spec/MSnSet.utils')

library(ggplot2)
library(ggfortify)
library(cowplot)

source('loadHumanPancData.R')

 metadata <- isletMeta%>%
   select(IsletStatus,IsletOrNot,Plex,`Grid Number`)%>%
   tibble::rownames_to_column('Spot')
 
 normtab <- do.call(rbind,lapply(crosstabList,function(x){
   newx<-x%>%
   tidyr::separate(feature,into=c('id','upid','protein'),
                   sep='\\|')%>%
   dplyr::select(-c(id,upid))
   #ibble::column_to_rownames('protein')
   newx<-newx%>%
     tidyr::pivot_longer(2:ncol(newx),names_to='Spot', values_to='logRatio')%>%
     tidyr::separate(Spot,into=c('Image','S','Xcoord','Ycoord'),sep='_',remove=F)%>%
     dplyr::select(-S)
   newx
   })
)%>%
  left_join(metadata)
 
 
 fulltab <- do.call(rbind,lapply(crosstabList2,function(x){
   newx<-x%>%
   tidyr::separate(feature,into=c('id','upid','protein'),
                   sep='\\|')%>%
   dplyr::select(-c(id,upid))
   #ibble::column_to_rownames('protein')
   newx<-newx%>%
     tidyr::pivot_longer(2:ncol(newx),names_to='Spot', values_to='logRatio')%>%
     tidyr::separate(Spot,into=c('Image','S','Xcoord','Ycoord'),sep='_',remove=F)%>%
     dplyr::select(-S)
   newx
   })
)%>%
  left_join(metadata)
 
 ##now save the fulltab to supp data table 1
 
write.table(fulltab,'suppTable1.csv',sep=',',quote=F,row.names=F)
```

Now we have a sense of how the data are distributed. Obviously this varies based on the normalization that we use in the original data.


## BayesSpace

First format as `SingleCellExperiment` to make sure that the data can be fed in to BayesSpace. Add in in the known labels and coordinates.

```{r format data}
##now format to SingleCellExperiment
library(Matrix)

counts<-fulltab%>%
  dplyr::select(Spot,protein,logRatio)%>%
  tidyr::pivot_wider(values_from=logRatio,names_from=Spot)%>%
  tibble::column_to_rownames('protein')


norm.counts<-normtab%>%
  dplyr::select(Spot,protein,logRatio)%>%
  tidyr::pivot_wider(values_from=logRatio,names_from=Spot)%>%
  tibble::column_to_rownames('protein')

colData <- fulltab%>%
    dplyr::select(Spot,Image,Xcoord,Ycoord,IsletStatus,IsletOrNot)%>%
  #distinct()%>%
#  tidyr::separate(Coords,into=c('S','col','row'),sep='_')%>%
  mutate(col=as.numeric(Xcoord),row=as.numeric(Ycoord))%>%
  distinct()

rowData <- data.frame(protein=rownames(counts))


p1<-ggplot(fulltab,aes(x=logRatio,fill=as.factor(`Grid Number`), shape=IsletStatus,alpha=0.5))+
  geom_density()+
  facet_grid(Image~.)+
  scale_fill_manual(values=RColorBrewer::brewer.pal(9,'Set1'))

p2<-ggplot(normtab,aes(x=logRatio,fill=as.factor(`Grid Number`), shape=IsletStatus,alpha=0.5))+
  geom_density()+
  facet_grid(Image~.)+
  scale_fill_manual(values=RColorBrewer::brewer.pal(9,'Set1'))

p3<-ggplot(fulltab,aes(x=logRatio,fill=IsletStatus,alpha=0.5))+geom_density()+facet_grid(Image~.)+scale_fill_manual(values=RColorBrewer::brewer.pal(9,'Set1'))

p4<-ggplot(normtab,aes(x=logRatio,fill=IsletStatus,alpha=0.5))+geom_density()+facet_grid(Image~.)+scale_fill_manual(values=RColorBrewer::brewer.pal(9,'Set1'))



  res=cowplot::plot_grid(p1,p3,p2,p4)
  print(res)
  
  ggsave('normalizedCountDistribution.pdf',res,height=8,width=11)

```

Now we have the data allegedly formatted as needed and can begin to infer things and can run through each spot individually.

## Cluster and plotting function

We put all the functionality into a single function for now.
Here we do preprocessing, which works, cluster analysis, the clustering. We iterate through 1-4 clusters, but generally always use 3 since that is what we are expecting. However, it seems that 4 also works.  

```{r spatial process, echo=FALSE}


plotDataPerImage<-function(image,countdat,prefix=''){


  #print(image)
  i1_spots<-subset(colData,Image==image)%>%
    tibble::column_to_rownames('Spot')
  
  cmat <- as.matrix(countdat[,rownames(i1_spots)])
  #print(dim(cmat))
  nas <- which(apply(cmat,1,function(x) all(is.na(x))))
  if(length(nas)>0)
    cmat <- cmat[-nas,]
  #print(dim(cmat))
  rowD <- rowData[rownames(cmat),]
  
  sce <- SingleCellExperiment(assays=list(logcounts=as(cmat, "dgCMatrix")),
                            rowData=rowD,
                            colData=i1_spots)

  set.seed(102)
  panc <- spatialPreprocess(sce, platform="ST", 
                              n.PCs=7, n.HVGs=1000, log.normalize=FALSE)

  panc <- qTune(panc, qs=seq(1, 4), platform="ST", d=5)
  qPlot(panc)
  
  panc2 <- spatialCluster(panc, q=2, platform="ST", d=5,
                           init.method="kmeans", model="normal", gamma=2,
                           nrep=2000, burn.in=200,
                           save.chain=TRUE)
  
  panc3 <- spatialCluster(panc, q=3, platform="ST", d=5,
                           init.method="kmeans", model="normal", gamma=2,
                           nrep=2000, burn.in=200,
                           save.chain=TRUE)
  
  panc4 <- spatialCluster(panc, q=4, platform="ST", d=5,
                           init.method="kmeans", model="normal", gamma=2,
                           nrep=2000, burn.in=200,
                           save.chain=TRUE)
  
    
  panc5 <- spatialCluster(panc, q=5, platform="ST", d=5,
                           init.method="kmeans", model="normal", gamma=2,
                           nrep=2000, burn.in=200,
                           save.chain=TRUE)
  
    panc6 <- spatialCluster(panc, q=6, platform="ST", d=5,
                           init.method="kmeans", model="normal", gamma=2,
                           nrep=2000, burn.in=200,
                           save.chain=TRUE)

#  pance2 <-spatialEnhance(panc2,q=2,d=5)  
#  pance3 <- spatialEnhance(panc3,q=3,d=5)
#  
#  pance4 <- spatialEnhance(panc4,q=4,d=5)

  
 cols<- RColorBrewer::brewer.pal(6,'Dark2')[c(1,2,3,6,4,5)]
    c0=clusterPlot(panc,
              palette=cols,
              label='IsletStatus')+
    ggtitle(paste('Image',image,'labels'))
    
    c1=clusterPlot(panc2,
              palette=cols)+
    ggtitle(paste('Image',image,'data'))
  
   c2=clusterPlot(panc3,
              palette=cols)+
    ggtitle(paste('Image',image,'data'))
  
  c3=clusterPlot(panc4,
              palette=cols)+
    ggtitle(paste('Image',image,'data'))
  
    c4=clusterPlot(panc5,
              palette=cols)+
    ggtitle(paste('Image',image,'data'))
    
      c5=clusterPlot(panc6,
              palette=cols)+
    ggtitle(paste('Image',image,'data'))
 # c4=clusterPlot(pance4,
#              palette=cols)+
#    ggtitle(paste('Image',image,'subspot'))

  res=cowplot::plot_grid(c0,c1,c2,c3,c4,c5)
  print(res)

  ggsave(paste0(prefix,'Islet',image,'.pdf'),res)
  ggsave(paste0(prefix,'Islet',image,'.png'),res)

  return(panc)
}


```

Now we can iterate over each of the spots. To see if they agree with our annotations. 

## Original images

```{r cluster, message=F,warning=F}

scObjs<-lapply(unique(colData$Image),function(x) plotDataPerImage(x,counts,'Un-normalized'))

##just do this manually

accuracy <- rbind(data.frame(image=c('0','10','1','2','3','4','7'),
                       TP=c(1,1,1,1,1,1,1),
                       FP=c(0,5,5,3,0,0,0),
                       FN=c(0,0,0,0,0,0,0),
                       TN=c(8,3,3,5,8,8,8),
                       numClusters=c(2,2,2,2,2,2,2)),
                  data.frame(image=c('0','10','1','2','3','4','7'),
                       TP=c(1,1,1,1,1,1,1),
                       FP=c(0,4,0,0,0,0,0),
                       FN=c(0,0,0,0,0,0,0),
                       TN=c(8,3,8,8,8,8,7),
                       numClusters=c(3,3,3,3,3,3,3)),
                  data.frame(image=c('0','10','1','2','3','4','7'),
                       TP=c(1,1,1,1,1,1,1),
                       FP=c(0,4,0,0,0,0,1),
                       FN=c(0,0,0,0,0,0,0),
                       TN=c(8,3,8,8,8,8,7),
                       numClusters=c(4,4,4,4,4,4,4)),
                   data.frame(image=c('0','10','1','2','3','4','7'),
                       TP=c(1,1,1,1,1,1,1),
                       FP=c(0,0,0,0,0,0,0),
                       FN=c(0,0,0,1,0,0,0),
                       TN=c(8,8,8,7,8,8,8),
                       numClusters=c(5,5,5,5,5,5,5)))%>%
  mutate(Specificity=TN/(TN+FP),Sensitivity=TP/(TP+FN))

ggplot(accuracy,aes(y=Specificity,fill=image,
                    x=as.factor(numClusters)))+
  geom_bar(stat='identity',position='dodge')+
  scale_fill_manual(values=RColorBrewer::brewer.pal(7,'Dark2'))                  

ggsave('bayesSpaceIsletId_unnorm.pdf')

```

this is the accuracy of the actual counts

## Normalized images

```{r norm}
scObjs<-lapply(unique(colData$Image),function(x) plotDataPerImage(x,norm.counts,'Normalized'))

norm.accuracy <- rbind(data.frame(image=c('0','10','1','2','3','4','7'),
                       TP=c(1,1,1,1,1,1,1),
                       FP=c(0,5,7,7,0,0,7),
                       FN=c(0,0,0,0,0,0,0),
                       TN=c(8,3,1,1,8,8,1),
                       numClusters=c(2,2,2,2,2,2,2)),
                  data.frame(image=c('0','10','1','2','3','4','7'),
                       TP=c(1,1,1,1,1,1,1),
                       FP=c(0,3,0,1,0,0,0),
                       FN=c(0,0,0,0,0,0,0),
                       TN=c(8,5,8,7,8,8,8),
                       numClusters=c(3,3,3,3,3,3,3)),
                  data.frame(image=c('0','10','1','2','3','4','7'),
                       TP=c(1,1,1,1,1,1,1),
                       FP=c(0,2,3,1,0,0,1),
                       FN=c(0,0,0,0,0,0,0),
                       TN=c(8,6,5,7,8,8,7),
                       numClusters=c(4,4,4,4,4,4,4)),
                   data.frame(image=c('0','10','1','2','3','4','7'),
                       TP=c(1,1,1,1,1,1,1),
                       FP=c(0,2,0,1,0,0,0),
                       FN=c(0,0,0,0,0,0,0),
                       TN=c(8,6,8,7,8,8,8),
                       numClusters=c(5,5,5,5,5,5,5)))%>%
  mutate(Specificity=TN/(TN+FP),
         Sensitivity=TP/(TP+FN))



ggplot(norm.accuracy,aes(y=Specificity,fill=image,
                    x=as.factor(numClusters)))+
  geom_bar(stat='identity',position='dodge')+
  scale_fill_manual(values=RColorBrewer::brewer.pal(7,'Dark2'))                  

ggsave('bayesSpaceIsletId_norm.pdf')
```

To some extent they do, but in other cases they do not. We need to check annotations to ensure they are correct. How can we test to see if they are in agreement? 

