---
title: "HubMap pancreata exploratory figures and variance analysis"
output: html_document
---

This document describes the basic analysis we need to showcase the value of the Hubmap data and proteomics analysis.

```{r loading, echo=F, warning=F, message=F}
library(dplyr)
library(purrr)
library(devtools)
#library(MSnSet.utils)
if(!require(MSnSet.utils))
  devtools::install_github('PNNL-Comp-Mass-Spec/MSnSet.utils')

library(ggplot2)
library(ggfortify)
library(cowplot)

source('loadHumanPancData.R')

```

Note that we store all files on synapse at [http://synapse.org/hubmap](http://synapse.org/hubmap). This includes basic data files and analysis results.

Now we have a few ways to normalize.

### Supplemental analysis to test normalization

We wanted to confirm that there was no need to filter the data or zero it out. Once we scale the data via PCA we get decent sample clustering.

```{r PCA plots}

#now we deal with the larger crosstab matrix

red.crosstab <- crosstab[rowMeans(!is.na(crosstab)) >= .5, ]
red.crosstab[which(is.na(red.crosstab),arr.ind=TRUE)]<-0.0

nz.crosstab<-crosstab
nz.crosstab[which(is.na(crosstab),arr.ind=TRUE)]<-0.0
#crosstab <- crosstab[rowMeans(!is.na(crosstab)) >= .5, ]

pc.z<-prcomp(t(nz.crosstab), center=TRUE,scale.=TRUE)
pc.red<-prcomp(t(red.crosstab), center=TRUE,scale.=TRUE)


#autoplot(pc,data=isletMeta,colour='IsletNumber')

p1<-autoplot(pc.z,data=isletMeta,colour='IsletStatus')+ggtitle('Full data')
p2<-autoplot(pc.red,data=isletMeta,colour='IsletStatus')+ggtitle("Reduced data")
cowplot::plot_grid(p1,p2)

##now plotting  with msn code
isletMeta <- isletMeta[colnames(crosstab), ]

##we keep NA values in m1 and m2
red.crosstab <- crosstab[rowMeans(!is.na(crosstab)) >= .5, ]

m1 <- MSnSet(exprs = red.crosstab, pData = isletMeta)

m2<- MSnSet(exprs = crosstab, pData = isletMeta)
##Need to find the plot_pca funciton, is it in msnbase.utils?
#plot_pca(m1, phenotype = "Islet_Number") #no batch effect
p3=plot_pca(m1, phenotype = "IsletStatus",z_score=T)+ggtitle('Reduced data') #check biplot = T?
p4=plot_pca(m2, phenotype = "IsletStatus",z_score=T)+ggtitle("Full data") #check biplot = T?

cowplot::plot_grid(p3,p4)

p5=plot_pca(m1, phenotype = "IsletNumber",z_score=T)+ggtitle('Reduced data') #check biplot = T?
p6=plot_pca(m2, phenotype = "IsletNumber",z_score=T)+ggtitle("Full data") #check biplot = T?

cowplot::plot_grid(p5,p6)
#save(m1, file = "combinedIslets_Results_New/combinedMSnSet.RData", compress = T)


```


Now that we can summarize the overlap between various regions we can also evaluate how various proteins vary. Basically there are over 1000 proteins that are generally up-regulated and down-regulated in islet. 

```{r variable proteins}

library(leapR)

#data("ncipid")
 data('krbpaths')
##map features to Gene Names
map<-read.table('uniprotMap.txt',header = TRUE)


var.vals <- apply(exprs(m2),1,var,na.rm=T)%>%sort()%>%
as.data.frame()%>%
tibble::rownames_to_column('feature')%>%
  tidyr::separate(feature,sep='\\|',into=c('sp','id','From'))%>%
  left_join(map)%>%
  dplyr::select(-c(sp,id,From))%>%
  #subset(!is.na(var.vals))%>%
  subset(!is.na(To))%>%
  distinct()

colnames(var.vals)[1]<-'variance'
#rownames(var.df)
hist(log10(var.vals$variance))



##now remove dupes
dupes<-var.vals$To[which(duplicated(var.vals$To))]

non.dupes<-var.vals%>%
  subset(!To%in%dupes)

fixed.dupes<-var.vals%>%
    subset(To%in%dupes)%>%
    group_by(To,.drop=F)%>%
    summarize(maxVar=max(variance))%>%
    left_join(var.vals)%>%
    subset(maxVar==variance)%>%
  dplyr::select(-maxVar)

full.var<-rbind(non.dupes,fixed.dupes)%>%
  tibble::remove_rownames()%>%
  tibble::column_to_rownames('To')%>%
  arrange(desc(variance))

##Now we can do analysis on thoes that are changing or not changing.
order.res<-leapR(datamatrix=full.var,krbpaths,'enrichment_in_order',primary_columns='variance',minsize=10)%>%
  subset(BH_pvalue<0.01)%>%
 # dplyr::select(ingroup_n,ingroup_mean,BH_pvalue,zscore)%>%
  arrange(BH_pvalue)

p<-plotLeapR(order.res,'REACTOME')

ggsave('mostVariableReactome.pdf',width=10)

```


## Lastly lets plot some single proteins to see what they look like

Insulin, glucagon, etc.

```{r single prots}

#prots<-c('INS_HUMAN','GLUC_HUMAN')

#subset(fulltab,protein%in%prots)%>%ggplot(aes(x=IsletStatus,y=logRatio,fill=protein))+geom_boxplot()+facet_grid(Xcoord~Ycoord)

#ggsave('insGlucExpr.pdf')


#subset(normtab,protein%in%prots)%>%ggplot(aes(x=IsletStatus,y=logRatio,fill=protein))+geom_boxplot()+facet_grid(Xcoord~Ycoord)

#ggsave('insGlucNormExpr.pdf')

```

