---
title: "Distance based analysis"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(purrr)
library(devtools)


library(ggplot2)
library(ggfortify)
library(cowplot)
library(leapR)
source('loadHumanPancData.R')
```

## Map to distances
get data and metadata nad join to get distance data

```{r distnace data, message=FALSE, warning=FALSE, message=FALSE,error=FALSE}

#m2<- MSnSet(exprs = crosstab2, pData = isletMeta)

fixed.crosstab<-correctMissingProteins(fulltab)


long.tab<-fixed.crosstab%>%#exprs(m2)%>%
  as.data.frame()%>%
  tibble::rownames_to_column('protein')%>%
  tidyr::pivot_longer(cols=c(1:ncol(fixed.crosstab)+1),
                      names_to='sample',
                      values_to='logFC')%>%
  left_join(tibble::rownames_to_column(isletMeta,'sample'))%>%
  dplyr::select(protein,`sample`,logFC,`Distance to islet`,`Islet Number`)
  
ct<-function(x,y,method){
  ret=NA
  try(ret<-cor.test(x,y,method=method)$p.value)
  ret
}

cor.vals<-long.tab%>%
  dplyr::group_by(protein,`Islet Number`)%>%
  dplyr::summarize(corVal=cor(logFC,`Distance to islet`,use='pairwise.complete.obs',method='pearson'),
            corTest=ct(logFC,`Distance to islet`,method='pearson'))


  
cor.na.vals<-long.tab|>
  dplyr::group_by(protein,`Islet Number`)|>
  dplyr::summarize(corVal=cor(logFC,`Distance to islet`,use='pairwise.complete.obs',method='spearman'),
            corTest=ct(logFC,`Distance to islet`,method='spearman'))
write.table(cor.na.vals,file='suppTab6.csv',sep=',',quote=F,row.names=F,col.names=T)

##which ones are highly anti-correlated
```

## Gene mapping and enrichment

now we want to evaluate the enrichment of the proteins whose expression is correlated with distance (lower in islet) and anti-correlated (higher in islet) 

```{r mapping, warning=FALSE, message=FALSE}

map<-read.table('uniprotMap.txt',header = TRUE)
 
 # tidyr::separate(feature,sep='\\|',into=c('sp','id','From'))%>%
#  left_join(map)%>%

##TODO: summary table off differential expression with venn diagram
cor.names<-cor.na.vals%>%
#  tidyr::separate(protein,sep='',into=c('sp','id','From'))%>%
  dplyr::rename(From='protein')%>%
  left_join(map)%>%
  subset(!is.na(To))


cor.mat<-cor.names%>%
  dplyr::ungroup()%>%
  dplyr::select(-c(corTest,From))%>%
  dplyr::mutate(`Islet Number`=as.character(`Islet Number`))%>%
  tidyr::pivot_wider(names_from=`Islet Number`,values_from=corVal,values_fn=list(corVal=mean))#%>%
#  tibble::column_to_rownames('To')

##now remove dupes
if(any(duplicated(cor.mat$To))){
  dupes<-cor.mat$To[which(duplicated(cor.mat$To))]

  cor.mat<-cor.mat%>%
    subset(!To%in%dupes)%>%
    tibble::column_to_rownames('To')
}

hist(cor.names$corVal)

#autoplot(pc.vals,label=TRUE)
#ggsave('correlationPCA.pdf')
```

now we have all the correlation values

```{r go enrichment, message=F, warning=F}
library(leapR)
library(pheatmap)
#data("ncipid")

go.order.res<-lapply(unique(as.character(cor.names$`Islet Number`)),function(x) {
  print(x)
               leapR(datamatrix=as.data.frame(cor.mat),gosigs,'enrichment_in_order',
                     primary_columns=x,id_column='To')%>%
               subset(pvalue<0.01)%>%
  #  dplyr::select(ingroup_n,ingroup_mean,BH_pvalue,zscore)%>%
              mutate(image=x)%>%
               tibble::rownames_to_column('Pathway')
              })

go.order.tab<-do.call(rbind,go.order.res)%>%
  #subset(pvalue<0.05)%>%
  dplyr::select(Pathway,zscore,image)%>%
  mutate(Pathway=stringr::str_replace_all(Pathway," \\(GO:[0-9]*\\)",''))%>%
  tidyr::pivot_wider(names_from='image',values_from='zscore',values_fill=list(zscore=0.0))%>%
  tibble::column_to_rownames('Pathway')
  
terms<-which(apply(go.order.tab,1,function(x) length(which(x==0))<3))

pheatmap(go.order.tab[names(terms),],cellwidth = 10,cellheight=10,filename = 'goBPterms.pdf')
```

now lets do kegg
```{r KEGG enrichment, message=F, warning=F}

#data("ncipid")
data('krbpaths')


kegg.order.res<-lapply(unique(as.character(cor.names$`Islet Number`)), function(x) {
  print(x)
  leapR(datamatrix=as.data.frame(cor.mat),krbpaths,'enrichment_in_order',
                  primary_columns=x,id_column='To',minsize=5)%>%
    subset(pvalue<0.01)%>%
  #  dplyr::select(ingroup_n,ingroup_mean,BH_pvalue,zscore)%>%
    dplyr::mutate(image=x)%>%  
    tibble::rownames_to_column('pathway')
})

krb.order.tab<-do.call(rbind,kegg.order.res)%>%
  subset(BH_pvalue<0.1)%>%
  dplyr::select(zscore,image,pathway,pvalue,BH_pvalue,ingroupnames)

kegg.order.tab<-krb.order.tab[grep("KEGG",krb.order.tab$pathway),]
reac.order.tab<-krb.order.tab[grep("REAC",krb.order.tab$pathway),]

kegg.order.tab<-kegg.order.tab%>%
  dplyr::select(-c(pvalue,BH_pvalue,ingroupnames))%>%
  mutate(pathway=stringr::str_replace_all(pathway,'_',' '))%>%
  mutate(pathway=stringr::str_replace_all(pathway,'REACTOME|KEGG|BIOCARTA',''))%>%
  mutate(pathway=tolower(pathway))%>%
  tidyr::pivot_wider(names_from='image',values_from='zscore',values_fill=list(zscore=0.0))%>%
  tibble::column_to_rownames('pathway')
  
  
reac.order.tab<-reac.order.tab%>%
    dplyr::select(-c(pvalue,BH_pvalue,ingroupnames))%>%
  mutate(pathway=stringr::str_replace_all(pathway,'_',' '))%>%
  mutate(pathway=stringr::str_replace_all(pathway,'REACTOME|KEGG|BIOCARTA',''))%>%
  mutate(pathway=tolower(pathway))%>%
  tidyr::pivot_wider(names_from='image',values_from='zscore',values_fill=list(zscore=0.0))%>%
  tibble::column_to_rownames('pathway')
  

#terms<-which(apply(kegg.order.tab,1,function(x) length(which(x==0))<6))

pheatmap(kegg.order.tab)

pheatmap(kegg.order.tab,cellwidth = 10,cellheight=10,filename = 'keggterms.pdf')



terms<-which(apply(reac.order.tab,1,function(x) length(which(x==0))<4))

pheatmap(reac.order.tab[terms,])

pheatmap(reac.order.tab[terms,],cellwidth = 10,cellheight=10,filename = 'reacterms.pdf')

```

One of the reactome terms that was associated was insulin signaling, what if we focus on that?

```{r insu}

path<-subset(krb.order.tab,pathway=='REACTOME_INSULIN_SYNTHESIS_AND_SECRETION')%>%
  select(ingroupnames)
prots<-path[1,]%>%
  stringr::str_split(.,pattern=', ')%>%
  unlist()

uprots<-subset(map,To%in%prots)%>%
  dplyr::select("From")

insgluc<-fixed.crosstab[unlist(uprots),]%>%
  as.data.frame()%>%
  tibble::rownames_to_column('protein')%>%
  tidyr::pivot_longer(2:(1+ncol(fixed.crosstab)),names_to='sample',values_to='logRatio')%>%
  left_join(tibble::rownames_to_column(isletMeta,'sample'))

path.avg<-insgluc%>%
  dplyr::group_by(sample,`X-coord`,`Y-coord`,`Islet Number`,IsletOrNot)%>%
  summarize(meanExp=mean(logRatio))

##now we can plot grid
p<- ggplot(path.avg)+
   geom_raster(aes(x=`X-coord`,y=`Y-coord`,fill=meanExp))+
   scale_fill_viridis_c()+coord_equal()+
   geom_point(aes(x=`X-coord`,y=`Y-coord`,col=IsletOrNot,alpha=0.1))+
   scale_color_manual(values=c('black','white'))+
   facet_grid(.~`Islet Number`)+theme_classic()

p

##not great, just munges things together
with.signif<-cor.names%>%
  mutate(signif=(corTest<0.05))%>%
                   subset(From%in%unlist(uprots))

cor.means<-with.signif%>%
  ungroup()%>%
  dplyr::group_by(To)%>%
  dplyr::summarize(mval=mean(corVal))%>%
  arrange(mval)

p2<-with.signif%>%
  ggplot(aes(x=factor(To,levels=cor.means$To),y=corVal,col=signif))+
  geom_jitter()+
  theme(axis.text.x=element_blank())

p2

ggsave('allCorrelationsInsulinSig.pdf',p2)
  
  
```

Here we can filter out the correlations that are interesting within the insulin signaling pathway, and 
dive into specific proteins of interest

```{r proteins}

signif<-with.signif%>%
  subset(signif)%>%
  dplyr::group_by(From)%>%
  summarize(numSignif=n())%>%
  subset(numSignif>1)


insgluc<-fixed.crosstab[signif$From,]%>%
  as.data.frame()%>%
  tibble::rownames_to_column('From')%>%
  left_join(map)%>%
  dplyr::rename(protein='To')%>%
  tidyr::pivot_longer(2:(1+ncol(fixed.crosstab)),names_to='sample',
                      values_to='logRatio')%>%
  left_join(tibble::rownames_to_column(isletMeta,'sample'))

p<- ggplot(insgluc)+
   geom_raster(aes(x=`X-coord`,y=`Y-coord`,fill=logRatio))+
   scale_fill_viridis_c()+coord_equal()+
   geom_point(aes(x=`X-coord`,y=`Y-coord`,col=IsletOrNot,alpha=0.1))+
   scale_color_manual(values=c('black','white'))+
   facet_grid(protein~`Islet Number`)+theme_classic()

p

ggsave('sigInsulinSig.pdf',p,height=10)
```

these are the proteins within insulin signaling that re correalted with distance.

We can generalize this code now in a function

```{r all pathways}

#' this function iterates through all the correlations
#' and plots those with proteins with significant correlation
#' in a specific pathway
plotPathwayCors<-function(krb.order.tab,pathname){
  
  #print(pathname)

  path<-subset(krb.order.tab,pathway==pathname)%>%
    select(ingroupnames)

  prots<-path[1,]%>%
    stringr::str_split(.,pattern=', ')%>%
    unlist()

  uprots<-subset(map,To%in%prots)%>%
    dplyr::select("From")

  insgluc<-fixed.crosstab[unlist(uprots),]%>%
    as.data.frame()%>%
    tibble::rownames_to_column('protein')%>%
    tidyr::pivot_longer(2:(1+ncol(fixed.crosstab)),names_to='sample',values_to='logRatio')%>%
   left_join(tibble::rownames_to_column(isletMeta,'sample'))

  path.avg<-insgluc%>%
    dplyr::group_by(sample,`X-coord`,`Y-coord`,`Islet Number`,IsletOrNot)%>%
    summarize(meanExp=mean(logRatio))

  ##now we can plot grid
  p<- ggplot(path.avg)+
     geom_raster(aes(x=`X-coord`,y=`Y-coord`,fill=meanExp))+
     scale_fill_viridis_c()+coord_equal()+
    geom_point(aes(x=`X-coord`,y=`Y-coord`,col=IsletOrNot,alpha=0.1))+
    scale_color_manual(values=c('black','white'))+
     facet_grid(.~`Islet Number`)+theme_classic()

  

  ##nget proteins that are significantly correlated with distance
  with.signif<-cor.names%>%
    mutate(signif=(corTest<0.05))%>%
                     subset(From%in%unlist(uprots))

  cor.means<-with.signif%>%
    ungroup()%>%
    dplyr::group_by(To)%>%
    dplyr::summarize(mval=mean(corVal))%>%
    arrange(mval)

  p2<-with.signif%>%
    ggplot(aes(x=factor(To,levels=cor.means$To),y=corVal,col=signif))+
    geom_jitter()+
    theme(axis.text.x=element_blank())
  
#  p2

  ggsave(paste0('allCorrelations',pathname,'.pdf'),p2)
  
  signif<-with.signif%>%
    subset(signif)%>%
    dplyr::group_by(From)%>%
    summarize(numSignif=n())%>%
    subset(numSignif>1) ##number significant genes is at least 2
  
  #print(signif)
  if(nrow(signif)>1){
    insgluc<-fixed.crosstab[signif$From,]%>%
      as.data.frame()%>%
      tibble::rownames_to_column('From')%>%
      left_join(map)%>%
      dplyr::rename(protein='To')%>%
      tidyr::pivot_longer(2:(1+ncol(fixed.crosstab)),names_to='sample',
                          values_to='logRatio')%>%
      left_join(tibble::rownames_to_column(isletMeta,'sample'))
  
  p3<- ggplot(insgluc)+
     geom_raster(aes(x=`X-coord`,y=`Y-coord`,fill=logRatio))+
     scale_fill_viridis_c()+coord_equal()+
     geom_point(aes(x=`X-coord`,y=`Y-coord`,col=IsletOrNot,alpha=0.1))+
     scale_color_manual(values=c('black','white'))+
    facet_grid(protein~`Islet Number`)+theme_classic()

  #p4<-cowplot::plot_grid(p,p2,p3)

    ggsave(paste0('sig_',pathname,'.pdf'),p3,height=10)
  }
}

```


Here we can filter out those pathways that are significant across ore than 2 voxels, then plot

```{r splicing proteins}

sig<-krb.order.tab%>%
  group_by(pathway)%>%
  summarize(num=n())%>%subset(num>3) 
## significantly correlated in at least three images

lapply(sig$pathway,function(x) plotPathwayCors(krb.order.tab,x))
```