---
title: "PCSF Tests"
author: "Sara Gosline"
date: "5/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(PCSF)
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Building trees out of voxels

The idea behind this approach is to build steiner trees out of individual voxels. In doing so we can make the voxels more functionally relevant. 

```{r cars}

syn=reticulate::import('synapseclient')
sync=syn$login()

tablist<-sync$tableQuery('select * from syn30806277')$asDataFrame()%>%
  dplyr::select(id,islet)

metadata <- sync$tableQuery('select * from syn30070462')$asDataFrame()%>%
  mutate(s='S')%>%
  unite(Islet_Number,s,Xcoord,Ycoord,sep='_',col='spot')

up_map <- read.table("uniprotMap.txt",col.names=c('protein','genename'))
fulltab<-do.call(rbind,apply(tablist,1,function(x)
  read.table(sync$get(x[['id']])$path)%>%
    mutate(islet=x[['islet']])))%>%
  tibble::rownames_to_column('geneid')%>%
  tidyr::separate(geneid,into=c('id','upid','protein'),
                  sep='\\|')%>%
  tidyr::pivot_longer(starts_with('S_'),
                      names_to='Coords',
                      values_to='logRatio')%>%
  unite(islet,Coords,sep='_',col='spot',remove = FALSE)%>%
  left_join(metadata)%>%
  left_join(up_map)

```

## PCA of each voxel

How do the voxels compare to each other? 


```{r pressure, echo=FALSE}
library(ggfortify)

df<-fulltab%>%dplyr::select(spot,logRatio,genename)%>%
  distinct()%>%
  subset(genename!='')%>%
  tidyr::pivot_wider(names_from=spot,values_from = logRatio,
                     values_fn = list(logRatio=mean))%>%
  tibble::column_to_rownames('genename')

mtab<-fulltab%>%dplyr::select(spot,islet,Coords,Plex,Islet_status)%>%
  distinct()%>%
  tibble::column_to_rownames('spot')

nas<-which(apply(df,1,function(x) any(is.na(x))))

pca_res <- prcomp(t(df[-nas,]),scale.=TRUE,center = TRUE, na.action=na.omit)

##does not work yet
p<-autoplot(pca_res,data=mtab,colour='Islet_status')
print(p)
```


## Get most expressed proteins in each voxel

First we bracket the data into the top vs bottom proteins to see how they are expressed.


```{r get top proteins}

##group by islet, take top 2.5 and bottom 2.5% of proteins
  
limits <- fulltab%>%
  dplyr::select(spot,logRatio)%>%
  dplyr::group_by(spot)%>%
  summarize(top=quantile(logRatio,0.975,na.rm=T),bottom=quantile(logRatio,0.025,na.rm=T))

##now determine if proteins are in top or bottom or none
topTerms <- fulltab%>%left_join(limits)%>%
  mutate(isTop=logRatio>top)%>%
  mutate(isBottom=logRatio<bottom)%>%
  subset(isTop|isBottom)

##how many top proteins in each spot? plot this
ggplot(topTerms)+geom_jitter(aes(x=islet,y=logRatio,col=Islet_status))


```

##Now run networks on top prots

iterate over the top proteins in each voxel 

```{r run net, echo=FALSE, warning=FALSE,message=FALSE}


library(PCSF)
data('STRING')
runPcsfOnProts<-function(tab,spot){
  terms<-tab$logRatio
  spot=unlist(spot)
#  print(spot)
  gene<-tab$genename
  ##map protein to gene name
  names(terms)<-gene
  ppi <- construct_interactome(STRING)
  
  ##return gene names, prizes
  print(spot)
  nres<-data.frame(nodes=c(""),prizes=c(""),spot=spot)
  res<-NULL
  try(res<-PCSF::PCSF(ppi=ppi,terminals = terms))
  if(!is.null(res) && length(vertex.attributes(res)$name)>1)
    nres<-data.frame(nodes=vertex.attributes(res)$name,
                     prizes=vertex.attributes(res)$prize,
                     spot=spot)
  return(nres)
}

##build network for each voxel

#now build networks based on top terms alone 
nets<-topTerms%>%
  subset(isTop)%>%
  dplyr::select(spot,genename,logRatio)%>%
 # separate(protein,into=c('gene','human'),sep='_')%>%
#  dplyr::select(spot,genename,logRatio)%>%
  group_by(spot)%>%
  dplyr::group_map(.f=runPcsfOnProts,.keep=FALSE)%>%
  do.call(rbind,.)

##now join with data from original table
##map distance between each network?
toplot<-fulltab%>%dplyr::select(genename,spot,logRatio,Islet_status)%>%
  #tidyr::separate(protein,into=c('nodes','hum'),sep='_')%>%
  dplyr::rename(nodes='genename')%>%
  right_join(nets,by=c('nodes','spot'))
##for each list of proteins in each voxel network, identify overlap with expresion data

```



## then plot PCA of new reduced expression set
```{r pca vals}

##PCA of prizes
df<-toplot%>%dplyr::select(spot,prizes,nodes)%>%
  mutate(prizes=as.numeric(prizes))%>%
  distinct()%>%
  tidyr::pivot_wider(names_from=spot,values_from = prizes,
                     values_fill = list(prizes=0.0),
                     values_fn = list(prizes=mean))%>%
  tibble::column_to_rownames('nodes')

mtab<-toplot%>%
  dplyr::select(spot,Islet_status)%>%
  distinct()%>%
  subset(!is.na(Islet_status))%>%
  tibble::column_to_rownames('spot')

nac<-which(colMeans(df,na.rm=T)!=0)
nar<-which(rowMeans(df,na.rm=T)!=0)
pca_res <- prcomp(t(df[nar,nac]),scale.=TRUE,center = TRUE)

p<-autoplot(pca_res,data=mtab,colour='Islet_status')
print(p)


##PCA of expression
df2<-toplot%>%dplyr::select(spot,logRatio,nodes)%>%
  mutate(logRatio=as.numeric(logRatio))%>%
  distinct()%>%
  replace_na(list(logRatio=0.0))%>%
  tidyr::pivot_wider(names_from=spot,
                     values_from = logRatio,
                     values_fn =list(logRatio=mean),
                     values_fill = list(logRatio=0.0))%>%
  tibble::column_to_rownames('nodes')

mtab2<-toplot%>%dplyr::select(spot,Islet_status)%>%
  distinct()%>%
  subset(!is.na(Islet_status))%>%
  tibble::column_to_rownames('spot')

nac<-which(colMeans(df2,na.rm=T)!=0)
nar<-which(rowMeans(df2,na.rm=T)!=0)
pca_res <- prcomp(t(df2[nar,nac]),scale.=TRUE,center = TRUE)

p<-autoplot(pca_res,data=mtab,colour='Islet_status')
print(p)

```


## Diffuse network in NetdifusR

```{R diffuse network}

##let's borrow a bit from PCSF
modstring <- STRING
modstring$cost <- 1-STRING$cost
igraph_ppi <- construct_interactome(modstring)

##https://github.com/USCCANA/netdiffuseR
#or 
###https://www.bioconductor.org/packages/devel/bioc/vignettes/diffuStats/inst/doc/diffuStats.pdf
```
## Now compare trees in PCA to see how they compare



## can we use tree output as input? 
