---
title: "BayesSpaceTest"
author: "Sara Gosline"
date: "5/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(BayesSpace)
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Format data
First grab data and format into single table. Evaluate distributions of values

```{r get data}

syn=reticulate::import('synapseclient')
sync=syn$login()

tablist<-sync$tableQuery('select * from syn30806277')$asDataFrame()%>%
  dplyr::select(id,islet)

metadata <- sync$tableQuery('select * from syn30070462')$asDataFrame()%>%
  mutate(s='S')%>%
  unite(Islet_Number,s,Xcoord,Ycoord,sep='_',col='spot')

fulltab<-do.call(rbind,apply(tablist,1,function(x)
  read.table(sync$get(x[['id']])$path)%>%
    mutate(islet=x[['islet']])))%>%
  tibble::rownames_to_column('geneid')%>%
  tidyr::separate(geneid,into=c('id','upid','protein'),
                  sep='\\|')%>%
  tidyr::pivot_longer(starts_with('S_'),
                      names_to='Coords',
                      values_to='logRatio')%>%
  unite(islet,Coords,sep='_',col='spot',remove = FALSE)%>%
  left_join(metadata)


```

Now we have a sense of how the data are distributed


## BayesSpace

First format as `SingleCellExperiment` to make sure that the data can be fed in to BayesSpace. Add in in the known labels and coordinates.

```{r format data}
##now format to SingleCellExperiment
library(Matrix)

counts<-fulltab%>%
  dplyr::select(spot,protein,logRatio)%>%
  tidyr::pivot_wider(values_from=logRatio,names_from=spot)%>%
  tibble::column_to_rownames('protein')

colData <- fulltab%>%
    dplyr::select(spot,islet,Coords,Islet_status)%>%
  distinct()%>%
  tidyr::separate(Coords,into=c('S','col','row'),sep='_')%>%
  mutate(col=as.numeric(col),row=as.numeric(row))
rowData <- data.frame(protein=rownames(counts))


ggplot(fulltab,aes(x=logRatio,fill=Coords,shape=Islet_status,alpha=0.5))+geom_density()+facet_grid(islet~.)+scale_fill_manual(values=RColorBrewer::brewer.pal(9,'Set1'))

ggplot(fulltab,aes(x=logRatio,fill=Islet_status,alpha=0.5))+geom_density()+facet_grid(islet~.)+scale_fill_manual(values=RColorBrewer::brewer.pal(9,'Set1'))
```

Now we have the data allegedly formatted as needed and can begin to infer things and can run through each spot individually.

## Cluster and plotting function

We put all the functionality into a single function for now.
Here we do preprocessing, which works, cluster analysis, the clustering. We iterate through 1-4 clusters, but generally always use 3 since that is what we are expecting. However, it seems that 4 also works.  

```{r spatial process, echo=FALSE}


plotDataPerImage<-function(image){

  library(RColorBrewer)
  library(cowplot)
  
  #print(image)
  i1_spots<-subset(colData,islet==image)%>%
    tibble::column_to_rownames('spot')
  
  cmat <- as.matrix(counts[,rownames(i1_spots)])
  #print(dim(cmat))
  nas <- which(apply(cmat,1,function(x) all(is.na(x))))
  if(length(nas)>0)
    cmat <- cmat[-nas,]
  #print(dim(cmat))
  rowData <- rowData[rownames(cmat),]
  sce <- SingleCellExperiment(assays=list(logcounts=as(cmat, "dgCMatrix")),
                            rowData=rowData,
                            colData=i1_spots)

  set.seed(102)
  panc <- spatialPreprocess(sce, platform="ST", 
                              n.PCs=7, n.HVGs=1000, log.normalize=FALSE)

  panc <- qTune(panc, qs=seq(1, 4), platform="ST", d=5)
  qPlot(panc)
  
  panc3 <- spatialCluster(panc, q=3, platform="ST", d=5,
                           init.method="kmeans", model="normal", gamma=2,
                           nrep=2000, burn.in=200,
                           save.chain=TRUE)
  
  panc4 <- spatialCluster(panc, q=4, platform="ST", d=5,
                           init.method="kmeans", model="normal", gamma=2,
                           nrep=2000, burn.in=200,
                           save.chain=TRUE)
  
  pance3 <- spatialEnhance(panc3,q=3,d=5)
  
  pance4 <- spatialEnhance(panc4,q=4,d=5)

  
 cols<- RColorBrewer::brewer.pal(6,'Dark2')[c(1,2,3,6)]
    c0=clusterPlot(panc,
              palette=cols,
              label='Islet_status')+
    ggtitle(paste('Image',image,'labels'))
    
    c1=clusterPlot(panc3,
              palette=cols)+
    ggtitle(paste('Image',image,'data'))
  
   c2=clusterPlot(panc4,
              palette=cols)+
    ggtitle(paste('Image',image,'data'))
  
  c3=clusterPlot(pance3,
              palette=cols)+
    ggtitle(paste('Image',image,'subspot'))
  
  c4=clusterPlot(pance4,
              palette=cols)+
    ggtitle(paste('Image',image,'subspot'))

  res=cowplot::plot_grid(c0,c1,c2,c3,c4)
  print(res)

  ggsave(paste0('Islet',image,'.pdf'),res)
  ggsave(paste0('Islet',image,'.png'),res)

  return(panc)
}

```

Now we can iterate over each of the spots. To see if they agree with our annotations. 

```{r cluster}

scObjs<-lapply(unique(colData$islet),function(x) plotDataPerImage(x))

```

To some extent they do, but in other cases they do not. We need to check annotations to ensure they are correct.
